Blum axioms and nondeterministic computation of functions

### 2016-04-13 paper.tex
paper.tex: Mention complexity classes and the Union Theorem.

Before extending these complexity measures to nondeterministic machines,
we must first provide an acceptable Gödel numbering.

>>>>

\begin{definition}
    Given a complexity measure $\Phi$ for an acceptable Gödel numbering $\phi$
    and a total recursive function $f : \mathbb N \to \mathbb N$,
    define the \emph{complexity class} $\mathcal C_\Phi(f)$ by
    (\cite[p.~232]{Kozen2006})
    \begin{equation*}
        \mathcal C_\Phi(f) = \{
            \phi_i \mid \Phi_i(x) \leq f(|x|)\ \text{almost everywhere\footnotemark}
        \}.
    \end{equation*}
    \footnotetext{
        \emph{Almost everywhere} means that the inequality $\Phi_i(x) \leq f(|x|)$
        holds for all but a finite number of different $x$.
    }
\end{definition}

This formalizes the notion of complexity class.
There are some important theorems concerning complexity classes
under the Blum axioms;
we mention only the Union Theorem \cite[p.~234]{Kozen2006}.
It states that,
if $\{f_i\}$ is any recursive list of increasing functions
such that $f_i(n) \leq f_{i+1}(n)$ for all $x$
(that is, each $f_i$ is greater than the previous),
then there is some function $g$ such that
\begin{equation*}
    \mathcal C_\Phi(g) = \bigcup_{i \in \mathbb N} \mathcal C_\Phi(f_i).
\end{equation*}
That is, the class $\mathcal C_\Phi(g)$ contains \emph{exactly}
all functions present in the classes $\mathcal C_\Phi(f_i)$.
Choosing $\Phi$ as the time complexity and $f_i(n) = n^i$
(the polynomial functions),
the union in the right is exactly the class $\P$,
the problems solvable in polynomial time.
By the Union Theorem, $\P$ is $\mathcal C_\Phi(g)$
for some recursive function $g$,
so even though $\P$ does not have an easily specifiable bounding function
such function exists nevertheless.
